Code for Breast Cancer Pain Study

### 2020-08-10 manuscript_code_20200809.Rmd
table updated

```{r}
final.data <- 
  cdmd.pvt %>%
  select("Pain",
         "Poverty","Blight",
         "Occupation","Marital.status","Insurance.type",
         "Cancer.stage","Chemo","Visits",
         "Race","Age",
         "Metastasis","Histology",
         "Primary.Site.1","Primary.Site.2",
         "HER2","ER","PR")
```

@@

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Visits",
             "Race","Age")

@@

```{r,echo=FALSE}
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Visits",
             "Race","Age")

get_summary = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  if (class(x) %in% c("numeric","integer")) {
    tmp.sum = t(t(c(paste0(round(mean(x),digits=2)," (",round(sd(x),digits=2),")"),
                    paste0(round(median(x),digits=2)," [",round(min(x),digits=2),",",round(max(x),digits=2),"]"))))
    rownames(tmp.sum) = paste(cname,c("Mean (SD)","Median [Min,Max]"),sep="_")
  } else {
    tmp.sum = t(t(summary(x)))
    tmp.sum[,1] = paste(tmp.sum[,1],paste0("(",round(100*(tmp.sum[,1]/sum(tmp.sum[,1])),digits=1),"%)"))
    rownames(tmp.sum) = paste(cname,rownames(tmp.sum),sep="_")
  }
  return(tmp.sum)
}
table.sum1 = do.call("rbind",sapply(1:ncol(final.data),get_summary))

@@

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Visits",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

```{r,echo=FALSE}
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Visits",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

```{r,echo=FALSE,eval=FALSE}
xlist = explvars[4:length(explvars)]

x = xlist[1]
p <- ggboxplot(final.data, x=x, y="Pain",remove="Unknown",
               color = x, palette="jco")
p + stat_compare_means()

x = xlist[2]
p <- ggboxplot(cdmd.pvt, x=x, y="Pain",remove="Unknown",
               color = x, palette="jco")
p + stat_compare_means()

x = xlist[3]
p <- ggboxplot(cdmd.pvt, x=x, y="Pain",remove="Unknown",
               color = x, palette="jco")
p + stat_compare_means()

x = xlist[4]
p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
               color = x, palette="jco")
p + stat_compare_means()

x = xlist[5]
p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
               color = x, palette="jco")
p + stat_compare_means()

x = xlist[6]
p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
               color = x, palette="jco")
p + stat_compare_means()

x = xlist[7]
p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
               color = x, palette="jco")
p + stat_compare_means()

x = xlist[8]
p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
               color = x, palette="jco")
p + stat_compare_means()

x = xlist[9]
p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
               color = x, palette="jco")
p + stat_compare_means()

@@

explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Visits",
               "Race","Age")
explvars.2 = c("Blight","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Visits",
               "Race","Age")
explvars = c("Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Visits",
             "Race","Age")
full.formula.1 = as.formula(paste(respvar," ~ ",paste(explvars.1,collapse=" + ")))
full.formula.2 = as.formula(paste(respvar," ~ ",paste(explvars.2,collapse=" + ")))
full.formula = as.formula(paste(respvar," ~ ",paste(explvars,collapse=" + ")))
```

@@

transform.p = function(x) {
  y = as.character(x)
  y[which(y==0)] = "<.001"
  return(y)
}
final.table$Unadj.P = transform.p(final.table$Unadj.P)
final.table$P.x = transform.p(final.table$P.x)
final.table$P.y = transform.p(final.table$P.y)
```

@@

table.count <-
  table.count %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.count)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.count)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.count)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.count)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.count)))) %>%
  # add_row(,.before=5+which(grepl(pattern="Visits",rownames(table.count)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.count)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.count))))
colnames(table.count) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

@@

table.zero <-
  table.zero %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.zero)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.zero)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.zero)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.zero)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.zero)))) %>%
  # add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.zero)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.zero)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.zero))))
colnames(table.zero) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

@@

@@

explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Visits",
               "Race","Age",
               "Metastasis","Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR")

explvars.2 = c("Blight","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Visits",
               "Race","Age",
               "Metastasis","Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR")

explvars = c("Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Visits",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

table.count <-
  table.count %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.count)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.count)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.count)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.count)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.count)))) %>%
  # add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.count)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.count)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.count)))) %>%
  add_row(,.before=8+which(grepl(pattern="Metastasis",rownames(table.count)))) %>%
  add_row(,.before=9+which(grepl(pattern="lobular",rownames(table.count)))) %>%
  add_row(,.before=10+which(grepl(pattern="Central",rownames(table.count)))) %>%
  add_row(,.before=11+which(grepl(pattern="Midline",rownames(table.count)))) %>%
  add_row(,.before=12+which(grepl(pattern="HER2Positive",rownames(table.count)))) %>%
  add_row(,.before=13+which(grepl(pattern="ERPositive",rownames(table.count)))) %>%
  add_row(,.before=14+which(grepl(pattern="PRPositive",rownames(table.count))))
colnames(table.count) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

@@

table.zero <-
  table.zero %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.zero)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.zero)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.zero)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.zero)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.zero)))) %>%
  # add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.zero)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.zero)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.zero)))) %>%
  add_row(,.before=8+which(grepl(pattern="Metastasis",rownames(table.zero)))) %>%
  add_row(,.before=9+which(grepl(pattern="lobular",rownames(table.zero)))) %>%
  add_row(,.before=10+which(grepl(pattern="Central",rownames(table.zero)))) %>%
  add_row(,.before=11+which(grepl(pattern="Midline",rownames(table.zero)))) %>%
  add_row(,.before=12+which(grepl(pattern="HER2Positive",rownames(table.zero)))) %>%
  add_row(,.before=13+which(grepl(pattern="ERPositive",rownames(table.zero)))) %>%
  add_row(,.before=14+which(grepl(pattern="PRPositive",rownames(table.zero))))
colnames(table.zero) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

>>>>

```{r}
final.data <- 
  cdmd.pvt %>%
  select("Pain",
         "Poverty","Blight",
         "Occupation","Marital.status","Insurance.type",
         "Cancer.stage","Chemo",
         "Race","Age",
         "Visits","Metastasis","Histology",
         "Primary.Site.1","Primary.Site.2",
         "HER2","ER","PR")
```

@@

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo",
             "Race","Age")

@@

```{r,echo=FALSE}
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo",
             "Race","Age")

get_summary = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  if (class(x) %in% c("numeric","integer")) {
    tmp.sum = t(t(c(paste0(round(mean(x),digits=2)," (",round(sd(x),digits=2),")"),
                    paste0(round(median(x),digits=2)," [",round(min(x),digits=2),",",round(max(x),digits=2),"]"))))
    rownames(tmp.sum) = paste(cname,c("Mean (SD)","Median [Min,Max]"),sep="_")
  } else {
    tmp.sum = t(t(summary(x)))
    tmp.sum[,1] = paste(tmp.sum[,1],paste0("(",round(100*(tmp.sum[,1]/sum(tmp.sum[,1])),digits=1),"%)"))
    rownames(tmp.sum) = paste(cname,rownames(tmp.sum),sep="_")
  }
  return(tmp.sum)
}
# table.sum1 = do.call("rbind",sapply(1:ncol(final.data),get_summary))

@@

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo",
             "Race","Age",
             "Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

```{r,echo=FALSE}
explvars = c("Pain","Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Visits","Metastasis",
             "Race","Age",
             "Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

@@

explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo",
               "Race","Age")
explvars.2 = c("Blight","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo",
               "Race","Age")
explvars = c("Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo",
             "Race","Age")
full.formula.1 = as.formula(paste(respvar," ~ ",paste(explvars.1,collapse=" + ")))
full.formula.2 = as.formula(paste(respvar," ~ ",paste(explvars.2,collapse=" + ")))
full.formula = as.formula(paste(respvar," ~ ",paste(explvars,collapse=" + ")))
```

@@

@@

table.count <-
  table.count %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.count)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.count)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.count)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.count)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.count)))) %>%
  # add_row(,.before=5+which(grepl(pattern="Metastasis",rownames(table.count)))) %>%
  add_row(,.before=5+which(grepl(pattern="Black",rownames(table.count)))) %>%
  add_row(,.before=6+which(grepl(pattern="Age",rownames(table.count))))
colnames(table.count) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

@@

table.zero <-
  table.zero %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.zero)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.zero)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.zero)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.zero)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.zero)))) %>%
  # add_row(,.before=5+which(grepl(pattern="Metastasis",rownames(table.zero)))) %>%
  add_row(,.before=5+which(grepl(pattern="Black",rownames(table.zero)))) %>%
  add_row(,.before=6+which(grepl(pattern="Age",rownames(table.zero))))
colnames(table.zero) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

@@

```{r,echo=FALSE,eval=FALSE}
# estimated probability and 95% CI:

get_estprob = function(estimates,x) {
  y = sum(estimates*x)
  return(exp(y)/(1+exp(y)))
}
## poverty using unadjusted model
tmp.table = final.unadj.list[[1]]

apply(log(tmp.table[grepl(pattern="zero",rownames(tmp.table)),1:3]),2,function(t) get_estprob(estimates=t,x=c(1,0)))

apply(log(tmp.table[grepl(pattern="zero",rownames(tmp.table)),1:3]),2,function(t) get_estprob(estimates=t,x=c(1,60)))

## insurance type using adjusted model
tmp.table = final.adj.1
tmp.table = tmp.table[grepl(pattern="zero",rownames(tmp.table)),]
x = c(1,mean(final.data$Poverty),0,0,0,0,0,0,0,0,1,0,1,0)
cbind(rownames(tmp.table),x)

apply(log(tmp.table[grepl(pattern="zero",rownames(tmp.table)),1:3]),2,function(t) get_estprob(estimates=t,x=c(1,mean(final.data$Poverty),0,0,0,0,0,0,0,0,0,0,0,0,0,0)))

apply(log(tmp.table[grepl(pattern="zero",rownames(tmp.table)),1:3]),2,function(t) get_estprob(estimates=t,x=c(1,mean(final.data$Poverty),0,0,0,0,0,0,1,0,0,0,0,0,0,0)))

table.zero = final.table[which(grepl("zero",rownames(final.table))),]
rownames(table.zero) = sapply(rownames(table.zero),function(x) unlist(strsplit(x,split="zero_"))[2]) # extract zero model

coef(final)[grepl(pattern="zero",names(coef(final)))]
```

@@

explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Visits","Metastasis",
               "Race","Age",
               "Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR")

explvars.2 = c("Blight","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Visits","Metastasis",
               "Race","Age",
               "Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR")

explvars = c("Poverty","Blight",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Visits","Metastasis",
             "Race","Age",
             "Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

@@

table.count <-
  table.count %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.count)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.count)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.count)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.count)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.count)))) %>%
  # add_row(,.before=5+which(grepl(pattern="Visits",rownames(table.count)))) %>%
  add_row(,.before=5+which(grepl(pattern="Metastasis",rownames(table.count)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.count)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.count)))) %>%
  # add_row(,.before=8+which(grepl(pattern="Metastasis",rownames(table.count)))) %>%
  add_row(,.before=8+which(grepl(pattern="lobular",rownames(table.count)))) %>%
  add_row(,.before=9+which(grepl(pattern="Central",rownames(table.count)))) %>%
  add_row(,.before=10+which(grepl(pattern="Midline",rownames(table.count)))) %>%
  add_row(,.before=11+which(grepl(pattern="HER2Positive",rownames(table.count)))) %>%
  add_row(,.before=12+which(grepl(pattern="ERPositive",rownames(table.count)))) %>%
  add_row(,.before=13+which(grepl(pattern="PRPositive",rownames(table.count))))
colnames(table.count) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

@@

table.zero <-
  table.zero %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.zero)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.zero)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.zero)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.zero)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.zero)))) %>%
  # add_row(,.before=5+which(grepl(pattern="Visits",rownames(table.zero)))) %>%
  add_row(,.before=5+which(grepl(pattern="Metastasis",rownames(table.zero)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.zero)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.zero)))) %>%
  # add_row(,.before=8+which(grepl(pattern="Metastasis",rownames(table.zero)))) %>%
  add_row(,.before=8+which(grepl(pattern="lobular",rownames(table.zero)))) %>%
  add_row(,.before=9+which(grepl(pattern="Central",rownames(table.zero)))) %>%
  add_row(,.before=10+which(grepl(pattern="Midline",rownames(table.zero)))) %>%
  add_row(,.before=11+which(grepl(pattern="HER2Positive",rownames(table.zero)))) %>%
  add_row(,.before=12+which(grepl(pattern="ERPositive",rownames(table.zero)))) %>%
  add_row(,.before=13+which(grepl(pattern="PRPositive",rownames(table.zero))))
colnames(table.zero) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")
